\chapter{Implementation and Evaluation}
\label{Chapter4}

This chapter describes how the design in Chapter~\ref{Chapter3} was implemented, the cloud deployment used to run and evaluate the system, the distribution of the mobile client, the security-hardening process applied to the platform, and an evaluation of the resulting system.

\section{Implementation Overview}

The backend services are implemented in Java 17. The Auth, Tracking, AI, and Dashboard services are built on Spring Boot~4.0 and organized as a Maven multi-module project so that shared code (for example, JWT validation logic) is written once and reused. The Therapist service is a separate Gradle project on Spring Boot~4.0, while the Social and Notification services are Gradle projects on the slightly earlier Spring Boot~3.3, reflecting when they were added to the system. Persistence throughout uses Spring Data JPA over PostgreSQL. The Social service's real-time chat is implemented with STOMP over WebSocket, and the Notification service integrates Firebase Cloud Messaging for push delivery in addition to email and an in-app inbox.

On the client side, the mobile application for teenagers is built with React Native~0.83 and React~19 in TypeScript, and the therapist-facing web dashboard is built with React~18, TypeScript, Vite, and Tailwind CSS.

\section{Deployment}

uMatter was originally deployed on Oracle Cloud Infrastructure (OCI). During the course of this thesis, the deployment was migrated in full to Microsoft Azure, under an Azure for Students subscription. The migration involved:

\begin{itemize}
\item Resolving Azure region-policy restrictions that initially blocked the target subscription from provisioning a virtual machine in the desired region.
\item Configuring Network Security Group (NSG) rules to open the ports required by the API gateway and by directly-exposed services (for example, RabbitMQ's management port during setup, and ports 80 and 443 for the reverse proxy).
\item Working around Azure Cloud Shell setup issues encountered while preparing the target VM.
\item Reproducing the auto-shutdown behavior available on the previous provider using Azure's built-in scheduled auto-shutdown, to control cost under the student subscription, with the containers configured to restart automatically on boot so they come back up automatically when the VM is next started.
\end{itemize}

The resulting deployment runs on a single, right-sized Azure VM hosting 20 containers: multiple Spring Boot JVM services, several PostgreSQL instances (one per owning service), RabbitMQ, Redis, MinIO, and pgAdmin. HTTPS routing is provided by Caddy as a reverse proxy, with DuckDNS supplying a stable hostname for the VM's dynamic address, and the VM's public IP configured as static so the hostname does not need to be re-pointed after the nightly auto-shutdown cycle. This mirrors the general topology described in Chapter~\ref{Chapter3}, with Caddy taking the reverse-proxy/HTTPS-termination role and Nginx continuing to serve as the internal API gateway between the proxy and the backend services.

\subsection{Mobile Application Distribution}

The mobile client for teenagers/patients was packaged and published to the Google Play Store, rather than distributed as a manually installed APK. Publishing through Google Play means end users install and update the app the same way as any other Android application, without needing to trust or manually side-load a build shared outside the store.

This distribution choice also shaped a concrete backend requirement: the published app communicates with the API gateway exclusively over HTTPS on port 443, because modern Android builds block plaintext HTTP traffic by default. Terminating TLS correctly at the reverse proxy, using a Let's Encrypt certificate obtained over port 80, was therefore not an optional hardening step but a hard prerequisite for the Play Store build to reach the backend at all. This is why ports 80 and 443 were treated as non-negotiable when the NSG rules were configured.

\section{Security Hardening}

Because uMatter handles sensitive personal data and coordinates real appointments, the implemented services were subjected to structured security audits organized around the OWASP Top~10 risk categories~\cite{2021-OWASP-Top10}, covering three areas in particular:

\begin{itemize}
\item \textbf{Business logic correctness}: verifying that workflows such as matching, booking, and permission granting behave correctly under normal and adversarial input, not only in the happy path.
\item \textbf{Access control}: specifically JPQL injection patterns in hand-written queries, insecure direct object references (for example, a user supplying another user's record identifier to an endpoint that does not re-check ownership), and mass assignment (a client supplying extra fields in a request body that get bound to fields the client should not be able to set, such as a role or a permission flag).
\item \textbf{Data integrity}: ensuring that concurrent requests cannot leave the system in an inconsistent state.
\end{itemize}

The most significant data-integrity issue identified was a race condition in appointment booking: two concurrent requests to book the same therapist for the same time slot could both pass an application-level availability check before either write was committed, resulting in a double booking. This was closed not by adding an application-level lock, which would not be safe against multiple service instances, but by making the slot claim a single atomic database operation: booking a slot is a conditional update that sets the slot's booked flag to true only where it is still false, so exactly one of two concurrent requests updates a row and the other is rejected. A unique constraint on the appointment's slot reference backs this up as a second line of defense. Because the guarantee lives in the database itself, it holds regardless of how many instances of the Therapist service are running concurrently.

The most significant access-control finding was in the AI service's own path to a user's tracking data, rather than in a caller manipulating another user's identifier. The Tracking service's internal context endpoint, used by the AI service to retrieve grounding data, was reachable by any authenticated internal caller. It performed no check that the caller had actually been granted access to that data, relying instead on network-level trust between services. This is precisely the class of implicit trust that the permission-based data-sharing model in Section~\ref{sec:datasec} is designed to remove. The fix followed the same pattern used elsewhere in the codebase: the AI companion was added as a reserved grantee in the existing access-grant table, and the context endpoint now checks for an active grant before returning anything, exactly as it already did for a therapist. No new authorization mechanism was introduced; an existing one was extended to a caller that had previously been exempted from it.

\section{Evaluation}

The system was evaluated along several dimensions: functional completeness against the design goals in Chapter~\ref{Chapter1}, security posture after hardening, feature positioning relative to the competitive landscape reviewed in Chapter~\ref{Chapter2}, and qualitative feedback from trial users and mental-health professionals, including the features they valued most and the wider findings that emerged from the design, trial, and consultation process.

\subsection{Functional Completeness}

All seven backend services described in Chapter~\ref{Chapter3} are implemented and deployed, and both client applications (mobile for teens, web for therapists) are functional against them. Table~\ref{tab:traceability} traces each requirement from Section~\ref{sec:requirements} to its implementation status.

Every requirement is met except FR8: all three notification channels (push, email, and the in-app inbox) work and fire end-to-end for appointment booking, but the other domain events intended to raise notifications are not currently wired to producers, so coverage across events is partial.

\begin{table}[htbp]
  \centering
  \caption{Implementation status of the requirements in Section~\ref{sec:requirements}}
  \label{tab:traceability}
  \begin{tabular}{|l|p{8.2cm}|l|}
    \hline
    \textbf{ID} & \textbf{Requirement (summary)} & \textbf{Status} \\
    \hline
    FR1 & Mood check-in, journal, mood calendar, and streaks & Met \\
    \hline
    FR2 & Sleep, nutrition, water, and breathing logs; automatic step count & Met \\
    \hline
    FR3 & Focus Mode check-in prompts, silent during Quiet Hours & Met \\
    \hline
    FR4 & AI companion grounded in tracking data only under an active grant & Met \\
    \hline
    FR5 & Guided breathing and meditation, Treasure Box, one-tap emergency area & Met \\
    \hline
    FR6 & Therapist discovery, booking, video consultation, and clinical notes & Met \\
    \hline
    FR7 & Friend and parent connections with real-time chat & Met \\
    \hline
    FR8 & Push, email, and in-app notifications from domain events & Partial \\
    \hline
    NFR1 & Privacy by default, grants enforced server-side & Met \\
    \hline
    NFR2 & Stateless JWT validation against a published JWKS key & Met \\
    \hline
    NFR3 & Integrity under concurrent booking & Met \\
    \hline
    NFR4 & Independent services on a single cloud VM & Met \\
    \hline
  \end{tabular}
\end{table}

\subsection{Security Posture}

The structured audits identified and closed concrete issues in each of the three areas examined (business logic, access control, and data integrity), including the appointment-booking race condition described above. This does not constitute a formal, independent penetration test, and should be read as an internal hardening pass rather than a certification of the system's security; further external review would be a reasonable next step before any production use involving real users.

\subsection{Comparison with Existing Platforms}

Revisiting the comparison in Table~\ref{tab:competitive}, the implemented system delivers on the feature combination that motivated this thesis: self-care tracking, an AI companion grounded in the user's own data, and peer social features, alongside licensed-therapist matching, booking, video consultation, and clinical notes, connected by a permission-based sharing model that none of the compared single-purpose platforms offer in combination. The clearest gap relative to mature commercial platforms such as BetterHelp and Talkspace~\cite{BetterHelp,Talkspace} is operational maturity: those platforms have been hardened and scaled in production for years, whereas uMatter, as a thesis project, has been validated on a single-VM deployment and an internal security-audit process rather than at production scale.

\subsection{Most-Valued Features}
\label{sec:most-valued}

Across the user trials and the three expert consultations conducted during this project (Section~\ref{sec:expert-findings}), two parts of uMatter stood out as the most consistently and highly valued: the \textbf{Treasure Box (Hộp Trân Quý)} and the suite of \textbf{daily reflective tracking} features. Both are described in the product tour in Appendix~\ref{...} % TODO: trỏ tới appendix/showcase phù hợp
and are singled out here because they attracted the strongest positive response.

\subsubsection{The Treasure Box (Hộp Trân Quý)}

% TODO: viết lại thành đoạn văn hoàn chỉnh. Ý chính:
\begin{itemize}
\item \textbf{What it is.} A personal store of positive ``core memories'' - photos, kind messages, small wins, life goals, and motivating reflections - that a user saves ahead of time and reopens whenever they need comforting.
\item \textbf{Why it was valued.} On a low day, negative bias makes good memories hard to recall; a pre-saved, reliable anchor back to the good gives users self-compassion they can reach for exactly when they need it.
\item \textbf{Where it came from.} The feature grew directly out of expert collaboration: Miss Pham Thi Thanh Qui first suggested a feature to reconnect users with their reasons for living and their motivation when feeling low, and mister Vuong Nguyen Toan Thien, the CEO of LUMOS, concretized that idea into the Treasure Box, grounding it in psychological research on positive memory and mood.
\item \textbf{How it fits.} It is one of the destinations in uMatter's ``psychology first-aid'' response to a negative mood, alongside guided breathing and the SOS crisis contacts.
% \item \textbf{[placeholder]} % TODO: chèn phản hồi/đánh giá cụ thể của người dùng thử (trích dẫn, số liệu nếu có).
\end{itemize}

\subsubsection{Daily Reflective Tracking}

% TODO: viết lại thành đoạn văn hoàn chỉnh. Ý chính:
\begin{itemize}
\item \textbf{What it is.} The everyday self-check-in habit at the heart of the app: mood check-ins, the emotional journal, sleep, nutrition and water, automatic step counting, and guided breathing - each surfaced as a friendly ``today'' card with weekly trends.
\item \textbf{Why it was valued.} Docter Pham Tran Thanh Nghiep assessed the daily reflective tracking features as ``already very good'' and already helpful to users' mental wellbeing. Miss Qui emphasised that physical and mental health are inseparable, endorsing the physical-health tracking (steps, sleep, nutrition, breathing) as legitimate mental-health signal.
\item \textbf{What makes it stick.} Gentle Focus-Mode reminders roughly every ninety minutes, daily achievement ``trophies'' for completing a chain of self-care tasks, and streaks turn tracking into a low-friction daily habit rather than a chore.
\item \textbf{Why it matters system-wide.} This daily data is also the foundation that grounds the AI companion, the therapist's view (under permission), and the trusted-friend view - so the feature most valued on its own is also what makes the rest of the platform work.
% \item \textbf{[placeholder]} % TODO: chèn phản hồi/đánh giá cụ thể của người dùng thử.
\end{itemize}

\subsection{Key Findings}
\label{sec:expert-findings}

Beyond the system itself, the design, trial, and consultation process surfaced a number of findings worth stating explicitly. % TODO: câu mở đầu, viết lại cho mượt.

\subsubsection{Findings from the Design and Development Process}

% TODO: viết lại. Ý chính (rút ra từ Chapter 3 & 4):
\begin{itemize}
\item \textbf{Unifying self-care and therapy is feasible behind one permission model.} A single microservices platform can serve both halves of the care journey without merging the data, as long as sharing is explicit, granular, and revocable.
\item \textbf{Network-level trust between services is not enough.} The most instructive security finding was that an internal caller (the AI service) could reach a user's tracking data with no explicit grant - implicit inter-service trust is exactly the class of assumption the permission model must remove, even for internal callers.
\item \textbf{Grounding the AI in the user's own data, with a safe fallback, is what makes it useful without over-reaching.} % TODO: mở rộng nếu muốn.
% \item \textbf{[placeholder]} % TODO: thêm finding kỹ thuật khác nếu cần.
\end{itemize}

\subsubsection{Findings from User Feedback}

% TODO: viết lại. Cần điền số người dùng thử / phương pháp khảo sát.
\begin{itemize}
\item \textbf{Daily journaling beats periodic surveys.} Trial users found the weekly/monthly surveys redundant next to the daily journal and 3-hourly mood check-ins; the surveys were dropped. Finding: high-frequency, low-friction daily logging yields richer and more meaningful data than periodic questionnaires.
\item \textbf{The most-loved features were the Treasure Box and daily tracking} (Section~\ref{sec:most-valued}),
which is why they are highlighted separately above.
% \item \textbf{[placeholder]} % TODO: điền các phản hồi khác của người dùng thử, số liệu đánh giá, trích dẫn. Mô tả cỡ mẫu và cách khảo sát (N người dùng thử, thời gian).
\end{itemize}

\subsubsection{Findings from Expert Consultations}

We consulted three mental-health professionals during the project: % TODO: có thể chuyển thành bảng.
\begin{itemize}
\item \textbf{Miss Pham Thi Thanh Qui} - counsellor, Psychological Counselling Office, University of Science, VNU-HCM (in person, 29 May 2026).
\item \textbf{Mister Vuong Nguyen Toan Thien} - CEO, LUMOS (online, 2 June 2026).
\item \textbf{Doctor Pham Tran Thanh Nghiep} - physician, Hoan My Sai Gon Hospital (online, 3 June 2026).
\end{itemize}

\noindent\textbf{What the experts validated and praised.} % TODO: viết lại thành văn.
\begin{itemize}
\item \textbf{Miss Qui} endorsed the option-based emotional journal, the gentle 3-hourly mood check-ins (for complete, meaningful data and for helping users pay attention to themselves), the 4-7-8 breathing exercise as a response to anger, the reward/trophy loop for completing a chain of self-care tasks, the revocable trusted-person sharing, and physical-health tracking such as steps.
\item \textbf{Mister Thien} contributed and praised the Treasure Box concept, the emotion timeline for complete mental-health data, the ``psychology first-aid'' bundle for negative moods (breathing, SOS contact, Treasure Box), and the SOS hotline page.
\item \textbf{Doctor Nghiep} gave the strongest endorsement of the daily reflective tracking features (``already very good'') and suggested adding curated meditation videos to the home screen for users to explore.
\end{itemize}

\noindent\textbf{Cross-cutting findings and design implications.} % TODO: viết lại thành văn.
\begin{itemize}
\item \textbf{Caution about AI in mental health.} All three experts were cautious about AI: two were broadly against relying on it, and one felt the AI should answer only very simple questions, refer harder ones to professionals, and not be given the user's private daily-tracking data. This validates uMatter's conservative AI design - grounded only under an explicit grant, with a safe ungrounded fallback and immediate crisis referral - and motivates the clinical review of the AI companion listed in Future Work.
\item \textbf{Caution about peer support.} Doctor Nghiep cautioned that friends may give advice contrary to professional guidance, suggesting clear in-app disclaimers and steering users toward trusted adults (e.g., parents) rather than peers. Design implication for the social layer. % TODO: nêu rõ đã/ sẽ xử lý thế nào.
\item \textbf{Iterating with experts changed the product.} Concrete features (Treasure Box, dropping weekly surveys, meditation videos, the psychology first-aid flow) came directly from these consultations, not from guesswork - itself a finding about how the app should be built.
\end{itemize}
